---
title: "riskimator"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{riskimator}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(riskimator)
```


# Cumulative Risk

The `causalRisk` package provides an estimator of the cumultative distribution of right-censored outcomes [@hubbard2000nonparametric]:

\[
\widehat{\Pr}(Y < t) = \frac{1}{n} \sum_{i = 1}^n \frac{\Delta_i I(Y_i < t)} { {\widehat{\Pr}} (\Delta_i = 1)}
\]

where $\Delta = I(Y < C)$ with a bit of algebra we can write this function as:

\[
h(t) = \frac{ \sum_i g(t)}{ \sum_i b }
\]

where:

\[
g(t) = a_1(t) \cdot a_2(t)
\]

\[
a_1(t) = I(Y_i < t)/1
\]

\[
a_2(t) = \Delta_i / \widehat{\Pr} (\Delta_i = 1)
\]

\[
b = 1
\]

The `monoidalEstimator` package provides a function `make_estimator` that creates the functions `g` and `h` above from a `list` of functions that create the `a` terms and a single `b` term.  Each term is created curried function that takes in `vs` (a `list` of data elements) and returns a `function` of whatever arguments are necessary (in this case `t`) and `...`. For this particular estimator, assuming the data are sorted by `min(Y, C)`, these could be written as:

